Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-06-15 mu/article/article.tex
Conclusions completed.

% Since its discovery, graphene has attracted a great interest in materials
% science and condensed matter physics being a new class of materials that are
% only one atom thick\cite{geimNM07}. 
% 
Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, high conductivity and strength \cite{geimNM07,
reinaNL08, novoselov2S07, balandinNL08} being then a perfect platform to be
used in two-dimensions electronic systems; however most electronic applications
are disabled by the absence of a semiconducting gap. Recent studies demonstrate
that the band gap of graphene can be opened by applying an electric
field,\cite{zhangN09} reducing the surface area,\cite{hanPRL07} or applying
uniaxial strain.\cite{niACSN08} Another possibility to open the gap is by
doping; this has been successfully achieved using nitrogen,\cite{weiNL2009}
boron-nitrogen,\cite{guoIJ11} silicon,\cite{colettiPRB10} noble-metals, and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene non centrosymmetric
structures both presenting a discernible band gap: the \emph{up} structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the \emph{alt} structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

@@

Spintronics devices are based on the control of spin and electrical charge
through electrical and optical methods.
% 
Semiconductor spintronics gives the development possibility of devices that can
perform high-volume information manipulation, processing, and storage, for
computing and communications technologies. \cite{awschalomNP2007}
% 
We have performed an \emph{ab initio} calculation for the SVI by one-photon
absorption of linearly polarized light in the \emph{up} and \emph{alt} 2D
hydrogenated graphene structures that could result in spintronics devices
development. This effect does not seem to have been reported previously.
% 
We calculated the responses for two cases: when the spin polarization is fixed
along the Cartesian coordinates and when the spin velocity is directed along
the $x$ and $y$ Cartesian directions on the $xy$ surface of the structures. We
found that  the SVI is very susceptible to the characteristics of the
structures presenting an anisotropic behavior.
% 
We found that for both structures it is possible to generate a SVI comparable
being the response of the \emph{alt} structure comparable with GaAs and CdSe
and the response of the \emph{up} structure much more intense than all.

@@

According to these results both structures are excellent candidates for
spintronics applications.

>>>>

% Since its discovery, graphene has attracted a great interest in materials
% science and condensed matter physics being a new class of materials that are
% only one atom thick\cite{geimNM07}. 
% 
Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08} being then a
perfect platform to be used in two-dimensions electronic systems; however most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals, and hydrogen.\cite{eliasS09,
guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene non centrosymmetric
structures both presenting a discernible band gap: the \emph{up} structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the \emph{alt} structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

@@

Spintronics devices are based on the control of spin and electrical charge
through electrical and optical methods.
% 
Semiconductor spintronics gives the development possibility of devices that can
perform high-volume information manipulation, processing, and storage, for
computing and communications technologies.\cite{awschalomNP2007} 
% 
We have performed an \emph{ab initio} calculation for the SVI by one-photon
absorption of linearly polarized light in the \emph{up} and \emph{alt} 2D
hydrogenated graphene structures that could result in spintronics devices
development. This effect does not seem to have been reported previously.
% 
We calculated the responses for two cases: when the spin polarization is fixed
along the Cartesian coordinates and when the spin velocity is directed along
the $x$ and $y$ Cartesian directions on the $xy$ surface of the structures. We
found that  the SVI is very susceptible to the characteristics of the
structures presenting an anisotropic behavior.
% 
We found that for both structures it is possible to generate a SVI comparable
being the response of the \emph{alt} structure comparable with GaAs and CdSe
and the response of the \emph{up} structure much more intense than all.

@@

Novel PSC devices for spintronics applications has been
proposed.\cite{gotteNat16, pershinPRB08} One of the difficulties to achieve the
development of spin current and PSC semiconductor devices is the fact that the
spin relaxation time in a semiconducting media is short disabling the spin
transport and then resulting in a no observable spin
current.\cite{murakamiSc03}
% 
The spin relaxation time in pure and doped graphene is long enough in the order
from nanoseconds to milliseconds. \cite{wojtaszekPRB13,ertler2009electron} The,
according to our results the \emph{alt} structure is a good candidate and the
\emph{up} structure an excellent candidate for the development of spintronics
devices that require PSC due to the high spin velocity transport.

\section{Acknowledgment} % (fold)

This work has been supported by \emph{Consejo Nacional de Ciencia y
Tecnolog\'ia} (CONACyT), M\'exico, Grant No. 153930.
